---
title: "Acomys Differential Expression Analysis Using VOOM (Merged & Unmerged)"
author: "Jason Orr Brant"
date: "`r Sys.time()`"
output: html_document
---

#### Set working directory and load required package

```{r echo=TRUE, results='hide', message=FALSE}
setwd("/Users/jasonbrant/Desktop/Dropbox (UFL)/RSEM_DATA/JULY_2016/ACO_PER_SE/")
x <- c("limma", "edgeR", "org.Mm.eg.db", "AnnotationDbi", "tidyr")
lapply(x, require, character.only = T)
```

#### Load count data and filter based on counts per million
Provide as input files containing raw counts; these were created using the Trinity/RSEM pipeline.
Change sample size and cpm based on each experiment.

```{r results='hold'}
rnaseqMatrix <- read.table("/Users/jasonbrant/Desktop/Dropbox (UFL)/RSEM_DATA/JULY_2016/ACO_PER_SE/aco_per_se_genes_counts_matrix.txt", header=T, row.names=1, com='')
rnaseqMatrix <- round(rnaseqMatrix)
dge <- DGEList(counts=rnaseqMatrix)
dim(dge)
keep <- rowSums(cpm(dge) > cpm(20,mean(dge$samples$lib.size))[1]) >= 4
dge <- dge[keep,]
dim(dge)
dge
```

#### Create treatment factor variable and set design 

```{r}
trt <- factor(c(1,1,1,1,2,2,2,2,3,3,3,3), labels = c("AcomysDay0", "AcomysDay7", "AcomysDay14"))
design <- model.matrix(~0 + trt)
```

#### Build overall model. 
Plot mean variance trend and MDS plots
```{r echo = TRUE, fig.show='hold'}
#png(filename = "./Acomys_mean_variance.png")
voom <- voom(dge, design, plot = T)
#dev.off()
#png(filename='./Acomys_MDS_plot.png')
plotMDS(voom, main="MDS plot")
#dev.off()
fit = eBayes(lmFit(voom,design))
top = topTable(fit,coef=ncol(design),number=Inf)
write.table(top, "./Acomys_DE_overall_model.txt", sep = '\t', quote = F, row.names = T)
```

#### AcomysDay7 vs AcomysDay0 pairwise contrasts

```{r fig.show='hold'}
contrast.matrix <- makeContrasts(contrasts = c("trtAcomysDay7-trtAcomysDay0"), levels = design)
fit2 <- contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)
#png(filename = "./AcomysDay7_vs_AcomysDay0_Q-Q_plot.png")
qqt(fit2$t, df=fit2$df.prior+fit2$df.residual,pch=16,cex=0.2,main="AcomysDay7 vs AcomysDay0 Q-Q plot")
abline(0,1)
#dev.off()
top=topTable(fit2, sort="none", n=Inf, coef=ncol(contrast.matrix))
source("/Users/jasonbrant/Desktop/Dropbox (UFL)/Regeneration_Project/RNA_Seq/ACOMYS/AGGREGATE/rnaseq_plot_funcs.R")
c = cpm(dge)
m = apply(c, 1, mean)
top2 = cbind(top, logCPM=log2(m[rownames(top)]))
#png("Acomys_genes_counts.counts.matrix.Day_7_vs_Day_0.voom.DE_results.MA_Plot.png")
plot_MA(top2$logCPM, top2$logFC, top2$'adj.P.Val')
#dev.off()
#png("Acomys_genes_counts.counts.matrix.Day_7_vs_Day_0.voom.DE_results.Volcano_Plot.png")
plot_Volcano(top2$logFC, top2$'adj.P.Val')
#dev.off()
top3 <- top2
top3$names <- rownames(top2)
rownames(top3) <- NULL
top3 <- top3[,c("names", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B", "logCPM")]
keys <- as.character(top3$names)
top4 <- select(org.Mm.eg.db, keys = keys, columns = "GENENAME", keytype = "SYMBOL")
resultsTable <- merge(top3, top4, by.x = "names", by.y = "SYMBOL", all.x = T)
write.table(resultsTable, "AcomysDay7_vs_AcomysDay0_MR_UN_RSEM_Voom_DE_ResultsTable.txt", row.names = F, quote = F, sep = '\t')
head(resultsTable)
```

#### AcomysDay14 vs AcomysDay0 pairwise contrasts

```{r fig.show='hold'}
contrast.matrix=makeContrasts(contrasts=c("trtAcomysDay14-trtAcomysDay0"), levels=design)
fit2=contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)
#png(filename='./AcomysDay14_vs_AcomysDay0_Q-Q_plot.png')
qqt(fit2$t , df=fit2$df.prior+fit2$df.residual,pch=16,cex=0.2,main="AcomysDay14 vs AcomysDay0 Q-Q plot")
abline(0,1)
#dev.off()
top=topTable(fit2, sort="none", n=Inf, coef=ncol(contrast.matrix))
source("/Users/jasonbrant/Desktop/Dropbox (UFL)/Regeneration_Project/RNA_Seq/ACOMYS/AGGREGATE/rnaseq_plot_funcs.R")
c = cpm(dge)
m = apply(c, 1, mean)
top2 = cbind(top, logCPM=log2(m[rownames(top)]))
#png("Acomys_genes_counts.counts.matrix.Day_14_vs_Day_0.voom.DE_results.MA_Plot.png")
plot_MA(top2$logCPM, top$logFC, top$'adj.P.Val')
#dev.off()
#png("Acomys_genes_counts.counts.matrix.Day_14_vs_Day_0.voom.DE_results.Volcano_Plot.png")
plot_Volcano(top$logFC, top$'adj.P.Val')
#dev.off()
top3 <- top2
top3$names <- rownames(top2)
rownames(top3) <- NULL
top3 <- top3[,c("names", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B", "logCPM")]
keys <- as.character(top3$names)
top4 <- select(org.Mm.eg.db, keys = keys, columns = "GENENAME", keytype = "SYMBOL")
resultsTable <- merge(top3, top4, by.x = "names", by.y = "SYMBOL", all.x = T)
write.table(resultsTable, "AcomysDay14_vs_AcomysDay0_MR_UN_RSEM_Voom_DE_ResultsTable.txt", row.names = F, quote = F, sep = '\t')
head(resultsTable)
```

#### AcomysDay14 vs AcomysDay7 pairwise contrasts

```{r fig.show='hold'}
contrast.matrix=makeContrasts(contrasts=c("trtAcomysDay14-trtAcomysDay7"), levels=design)
fit2=contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)
#png(filename='./AcomysDay14_vs_AcomysDay7_Q-Q_plot.png')
qqt(fit2$t , df=fit2$df.prior+fit2$df.residual,pch=16,cex=0.2,main="AcomysDay14 vs AcomysDay7 Q-Q plot")
abline(0,1)
#dev.off()
top=topTable(fit2, sort="none", n=Inf, coef=ncol(contrast.matrix))
source("/Users/jasonbrant/Desktop/Dropbox (UFL)/Regeneration_Project/RNA_Seq/ACOMYS/AGGREGATE/rnaseq_plot_funcs.R")
c = cpm(dge)
m = apply(c, 1, mean)
top2 = cbind(top, logCPM=log2(m[rownames(top)]))
#png("Acomys_genes_counts.counts.matrix.Day_14_vs_Day_7.voom.DE_results.MA_Plot.png")
plot_MA(top2$logCPM, top$logFC, top$'adj.P.Val')
#dev.off()
#png("Acomys_genes_counts.counts.matrix.Day_14_vs_Day_7.voom.DE_results.Volcano_Plot.png")
plot_Volcano(top$logFC, top$'adj.P.Val')
#dev.off()
top3 <- top2
top3$names <- rownames(top2)
rownames(top3) <- NULL
top3 <- top3[,c("names", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B", "logCPM")]
keys <- as.character(top3$names)
top4 <- select(org.Mm.eg.db, keys = keys, columns = "GENENAME", keytype = "SYMBOL")
resultsTable <- merge(top3, top4, by.x = "names", by.y = "SYMBOL", all.x = T)
write.table(resultsTable, "AcomysDay14_vs_AcomysDay7_MR_UN_RSEM_Voom_DE_ResultsTable.txt", row.names = F, quote = F, sep = '\t')
head(resultsTable)
```

#### Session Information

```{r}
Sys.time()
sessionInfo()
```
