---
title: "Voom Time Course"
author: "Jason Orr Brant"
date: "August 12, 2016"
output: html_document
---

#### Set working directory and load required package

```{r echo=TRUE, results='hide', message=FALSE}
setwd("~/Desktop/Dropbox (UFL)/RSEM_DATA/JULY_2016/TIME_COURSE/")
x <- c("limma", "edgeR", "org.Mm.eg.db", "AnnotationDbi", "tidyr")
lapply(x, require, character.only = T)
```

#### Load count data and filter based on counts per million
Provide as input files containing raw counts; these were created using the Trinity/RSEM pipeline.
Change sample size and cpm based on each experiment.

```{r results='hold'}
rnaseqMatrix <- read.table("~/Desktop/Dropbox (UFL)/RSEM_DATA/JULY_2016/TIME_COURSE/aco_and_mus_MR_gene_counts_matrix.txt", header=T, row.names=1, com='')
rnaseqMatrix <- round(rnaseqMatrix)
dge <- DGEList(counts=rnaseqMatrix)
dim(dge)
keep <- rowSums(cpm(dge) > cpm(20,mean(dge$samples$lib.size))[1]) >= 3
dge <- dge[keep,]
dim(dge)
dge
dge <- calcNormFactors(dge, method='TMM')
```

#### Create treatment factor variable and set design 

```{r}
trt <- factor(c(1,1,1,1,2,2,2,2,3,3,3,3,4,4,4,5,5,5,5,6,6,6,6), labels = c("AcomysDay0", "AcomysDay7", "AcomysDay14", "MusDay0", "MusDay7", "MusDay14"))
design <- model.matrix(~0 + trt)
```

#### Build overall model. 
Plot mean variance trend and MDS plots
```{r echo = TRUE, fig.show='hold'}
#png(filename = "./Acomys_vs_Mus_mean_variance.png")
voom <- voom(dge, design, plot = T) # normalize.method = "scale"
#dev.off()
#png(filename='./Acomys_vs_Mus_MDS_plot.png')
plotMDS(voom, main="MDS plot")
#dev.off()
fit = eBayes(lmFit(voom,design))
top = topTable(fit,coef=ncol(design),number=Inf)
write.table(top, "./Acomys_vs_Mus_DE_overall_model.txt", sep = '\t', quote = F, row.names = T)
```

# Determine which genes respond differently over time in Acomys relative to Mus (Days 0, 7 and 14)

```{r}
cont.dif <- makeContrasts(Dif_7day = (trtAcomysDay7-trtAcomysDay0)-(trtMusDay7-trtMusDay0),
    Dif14day=(trtAcomysDay14-trtAcomysDay7)-(trtMusDay14-trtMusDay7),levels=design)
fit2 <- contrasts.fit(fit, cont.dif)
fit2 <- eBayes(fit2)
#topTableF(fit2, adjust="BH")
top = topTableF(fit2, number=Inf)
top2 <- top
top2$names <- rownames(top)
keys <- as.character(top2$names)
top4 <- select(org.Mm.eg.db, keys = keys, columns = "GENENAME", keytype = "SYMBOL")
resultsTable <- merge(top2, top4, by.x = "names", by.y = "SYMBOL", all.x = T)

write.table(resultsTable, file = "./time_course_Mus_Aco_days0-7-14.txt", sep = '\t', quote = F, row.names = F)
```

# Determine which genes respond differently over time in Acomys relative to Mus (Days 0 and 7)

```{r}
cont.dif <- makeContrasts(Dif_7day = (trtAcomysDay7-trtAcomysDay0)-(trtMusDay7-trtMusDay0), levels=design)
fit2 <- contrasts.fit(fit, cont.dif)
fit2 <- eBayes(fit2)
#topTableF(fit2, adjust="BH")
top = topTableF(fit2, number=Inf)
top2 <- top
top2$names <- rownames(top)
keys <- as.character(top2$names)
top4 <- select(org.Mm.eg.db, keys = keys, columns = "GENENAME", keytype = "SYMBOL")
resultsTable <- merge(top2, top4, by.x = "names", by.y = "SYMBOL", all.x = T)

write.table(resultsTable, file = "./time_course_Mus_Aco_days0-7.txt", sep = '\t', quote = F, row.names = F)
```

# Determine which genes respond differently over time in Acomys relative to Mus (Days 0 and 14)

```{r}
cont.dif <- makeContrasts(Dif_14day = (trtAcomysDay14-trtAcomysDay0)-(trtMusDay14-trtMusDay0), levels=design)
fit2 <- contrasts.fit(fit, cont.dif)
fit2 <- eBayes(fit2)
#topTableF(fit2, adjust="BH")
top = topTableF(fit2, number=Inf)
top2 <- top
top2$names <- rownames(top)
keys <- as.character(top2$names)
top4 <- select(org.Mm.eg.db, keys = keys, columns = "GENENAME", keytype = "SYMBOL")
resultsTable <- merge(top2, top4, by.x = "names", by.y = "SYMBOL", all.x = T)

write.table(resultsTable, file = "./time_course_Mus_Aco_days0-14.txt", sep = '\t', quote = F, row.names = F)
```

# Determine which genes respond differently over time in Acomys relative to Mus (Days 7 and 14)

```{r}
cont.dif <- makeContrasts(Dif14day = (trtAcomysDay14-trtAcomysDay7)-(trtMusDay14-trtMusDay7), levels=design)
fit2 <- contrasts.fit(fit, cont.dif)
fit2 <- eBayes(fit2)
#topTableF(fit2, adjust="BH")
top = topTableF(fit2, number=Inf)
top2 <- top
top2$names <- rownames(top)
keys <- as.character(top2$names)
top4 <- select(org.Mm.eg.db, keys = keys, columns = "GENENAME", keytype = "SYMBOL")
resultsTable <- merge(top2, top4, by.x = "names", by.y = "SYMBOL", all.x = T)

write.table(resultsTable, file = "./time_course_Mus_Aco_days7-14.txt", sep = '\t', quote = F, row.names = F)
```

#### Session Information

```{r}
Sys.time()
sessionInfo()
```

